\subsection{Universality of the Saturation Scale}
  \label{subsec:universality}

  A central outcome of the analysis is the empirical stability of the
  saturation scale $a_\star$ when determined from the full SPARC sample
  (N = 175 galaxies).
  The scale is obtained through a global robust optimization,
  varying only the stellar disc mass-to-light ratio $\Upsilon_\star$
  on a galaxy-by-galaxy basis, while keeping $\Upsilon_{\mathrm bul}$ fixed.

  Using a median-based estimator and allowing
  $\Upsilon_\star \in [0.01,2.0]$,
  we obtain
  $a_\star \simeq 2.4 \times 10^3
  \,(\mathrm{km\,s^{-1}})^2\,\mathrm{kpc}^{-1}$.
  A trimmed-mean criterion yields
  $a_\star \simeq 2.8 \times 10^3
  \,(\mathrm{km\,s^{-1}})^2\,\mathrm{kpc}^{-1}$.
  The inferred scale is therefore stable at the $\sim 15\%$ level
  with respect to the choice of robust estimator.

  The resulting distribution of reduced $\chi^2$ values has a median
  $\chi^2_\nu \approx 2.7$, with approximately
  $60$--$65\%$ of galaxies satisfying $\chi^2_\nu < 5$
  and nearly $80\%$ satisfying $\chi^2_\nu < 10$.
  These values are obtained without introducing galaxy-dependent halo
  parameters or cosmological priors.

  The inferred scale remains within the same range when adopting
  plausible variations of the bulge mass-to-light ratio
  ($\Upsilon_{\mathrm bul} = 0.3$--$0.7$),
  indicating that $a_\star$ is not driven by fine-tuning of baryonic
  normalization.

  This supports the interpretation of $a_\star$ as a universal effective
  transition scale rather than as a phenomenological fitting parameter.
  All galaxy-to-galaxy variability is accounted for by the observed
  baryonic structure, with no halo-by-halo adjustment.

  In the present framework, the universality of the saturation scale
  reflects the existence of a bounded-response regime in the effective
  geometric description, rather than a modification of the fundamental
  gravitational law.
