\section{Galaxy Rotation Curves}
  \label{sec:rotation_curves}

  Throughout this section, the only galaxy-dependent free parameter
  is the stellar disc mass-to-light ratio $\Upsilon_\star$,
  varied within a physically motivated interval.
  The saturation scale $a_\star$ is determined once from the full sample
  and then held fixed.
  No additional halo parameters are introduced.

  We now test the effective model introduced in
  Section~\ref{sec:effective_model} against observed galaxy rotation curves.
  The goal of this section is not to achieve maximal fitting flexibility, but to
  assess whether a single effective saturation scale can reproduce the main
  kinematic features of disk galaxies using baryonic matter alone.

  A useful internal consistency requirement is that the acceleration scale
  governing the transition be determined directly from galactic kinematic data.
  For the SPARC sample, fitting the effective model to the observed rotation
  curves yields a characteristic transition scale of order
  $10^{-10}\,\mathrm{m\,s^{-2}}$.
  This value is obtained within the present framework solely from galaxy
  rotation data and does not rely on any cosmological prior or assumption
  regarding the Hubble constant.

  The numerical value is consistent with the empirical acceleration scale
  identified in MOND analyses~\cite{McGaugh2016}.
  In the present framework, however, this scale is not introduced as a
  fundamental modification of the gravitational law.
  It emerges as the transition point of a bounded-response regime in
  the effective kinematic description.

  This motivates treating $a_\star$ as a single universal effective scale
  consistently appearing in both high- and low-surface-density galaxies,
  without requiring modifications to early-time cosmology or additional
  dark components.

  \input{03-rotation_curves/001-data_method}
  \input{03-rotation_curves/002-fits}
  \input{03-rotation_curves/003-residuals}
  \input{03-rotation_curves/004-universality}
  \input{03-rotation_curves/005-rc_summary}
