\section{Conclusion}
  \label{sec:conclusion}

  We have examined whether two long-standing observational anomalies,
  namely flat galaxy rotation curves and the Hubble constant tension,
  may originate from a common effective mechanism operating in low-density environments.

  Within a minimal phenomenological framework characterized by a single
  saturation scale $a_\star$, we have shown that baryonic matter alone
  reproduces the main features of observed galaxy rotation curves across
  a wide range of systems, without halo-by-halo tuning and without
  introducing dark matter particles at the level of the effective description.
  This result is supported by global fits to the SPARC sample.

  The same saturation principle extends naturally to cosmological
  low-density regions.
  Cosmic voids probe the deepest saturation regime and provide a direct
  link between galactic low-acceleration phenomenology and late-time
  expansion inference.
  In such environments, the locally inferred expansion rate is predicted
  to deviate systematically from the global value inferred from
  early-universe observables, while the background Friedmann expansion
  remains unmodified.

  Crucially, the framework does not alter early-time physics and does not
  modify the global cosmological dynamics.
  Gravitational acceleration and the Hubble parameter are interpreted as
  effective inference rates emerging from a bounded-response structure,
  rather than as fundamental fields requiring modification.

  A distinctive strength of the proposal lies in its cross-sector
  consistency.
  The same empirically determined scale $a_\star$ governs both
  galactic rotation curves and environment-dependent expansion inference.
  Constraints obtained in one domain therefore directly restrict the
  allowed phenomenology in the other, preventing independent tuning of
  galactic and cosmological sectors.

  The framework is sharply falsifiable.
  It predicts measurable correlations between inferred values of $H_0$
  and large-scale density contrast, a redshift-dependent suppression of
  the local offset, and specific signatures in ultra-diffuse galaxy
  kinematics, redshift drift, and void weak lensing.
  Failure of these correlated predictions would invalidate the mechanism.

  The central claim of this work is therefore not that gravity is modified,
  nor that dark matter is replaced by a new fundamental constant,
  but that a bounded-response principle may underlie both galactic
  low-acceleration behavior and late-time expansion inference.
  Whether this effective description captures a genuine structural aspect
  of gravitational phenomenology or merely approximates deeper dynamics
  will be decided by forthcoming observational tests.

  Beyond this phenomenological core, Section~\ref{sec:history-branching-expansion} isolates a structural
  expansion sector with explicitly stratified statuses.
  Proved elsewhere and imported here: endpoint counting cannot carry an exponential component on the
  Heisenberg substrate, while projectively distinguishable histories follow the exact Pell count with rate
  $\log(1+\sqrt2)$~\cite{Beau2026tb}.
  Hypothesis: the rank--time dictionary $n=\log a$ (Hypothesis~\ref{hyp:H-dict}), under which the naive
  state-volume reading of expansion is quantitatively excluded.
  Structural, conditional on [H-ent-id]: the de Sitter limit as a calibrated spectral steady state
  (Section~\ref{subsec:desitter-steady-state}), with the numerical value of $H_\infty$ explicitly not
  derived (structural carrier yes, numerical value no).
  Conditional no-go: the extraction of an observable evolving equation of state from the exact full-history
  counts, under the dictionary hypothesis and minimal mappings~\cite{Beau2026tb}.
  Open: the entropy-to-power step of the large-angle capacity mechanism
  (Section~\ref{subsec:low-ell-prebranching}, developed in~\cite{Beau2026dhc}), and any evolving-$w$ route
  outside the exact counts.
  None of the phenomenological results of the preceding sections depends on these items.

\subsection*{Interpretive outlook}
  The following reading is interpretive: a structural interpretation of the results above, not itself a
  derived result.
  Within the programme, the picture this paper draws is of a universe whose late-time behaviour is governed
  by inference bounds rather than by new substances: rotation curves and the local expansion offset trace a
  bounded kinematic response in diffuse environments, while the accelerated component reflects the branching
  of projectively distinguishable histories --- expansion as a property of the projective record rather than
  of a dark-energy fluid.
  In this reading the asymptotic de Sitter phase is a driven steady state of the spectral-equilibrium
  sector, its entropy carrier and production rate fixed structurally, its single scale calibrated in the
  same sense as the induced Newton coupling.
  The same reading reshapes the earliest projectable epoch: the first admissibility ranks carry only a
  handful of distinguishable histories ($3$, $7$, $17$: the exact Pell count), so the beginning of the
  observable universe appears as a state of finite resolution rather than of infinite temperature.
  Temperature is then an effective parameter, meaningful only once the projective unfolding has produced
  enough distinguishable degrees of freedom to carry a statistical description; the hot early universe of
  the classical regime is untouched, while the extrapolation to an infinitely hot initial state is
  reclassified as an artefact of pushing the effective description beyond its domain of validity.
  A third party should situate these claims through their stratification: the phenomenological core is
  falsifiable now; the structural expansion sector is conditional on the stated dictionary and
  identification hypotheses; and no numerical prediction of $H_\infty$ is made.
