\begin{abstract}
  We investigate whether two long-standing observational anomalies---flat galaxy
  rotation curves and the Hubble constant tension---may originate from a common
  effective mechanism operating in low-density environments.

  We introduce a minimal phenomenological framework in which departures from
  Newtonian expectations arise from a saturation of the effective gravitational
  response at the level of kinematic inference, without invoking dark matter
  particles or altering early-universe cosmology.
  In the galactic regime, the well-known MOND phenomenology emerges as the
  low-density limit of this bounded-response framework, rather than as a
  fundamental modification of gravity.
  The framework reduces to standard Newtonian behavior in high-density regimes
  and introduces a smooth, bounded effective response in diffuse environments.

  Applied to galactic dynamics, this approach reproduces the main features of
  observed rotation curves across a wide range of galaxy types using baryonic
  matter alone, with stable parameter values and no halo-by-halo tuning.
  Explicit numerical comparisons with representative galaxies from the SPARC
  sample show that the effective saturation scale naturally correlates with
  observed surface-density-dependent trends.

  At cosmological scales, the same effective mechanism predicts
  environment-dependent deviations in the locally inferred expansion rate.
  For typical void density contrasts, an explicit order-of-magnitude estimate
  shows that the deep-saturation regime applies, yielding a fractional shift
  $\Delta H/H \simeq \delta_\rho f_{\max}$ without introducing any new
  cosmological degree of freedom.
  Cosmic voids therefore provide a natural setting for a systematic offset
  between local and global measurements of the Hubble constant, while
  preserving the background expansion history.

  We examine robustness, degeneracies with standard astrophysical effects, and
  limitations of the effective description, and we outline distinctive
  observational signatures, including environment-dependent redshift drift and
  weak lensing effects.
  The results suggest that galaxy rotation curves and the Hubble tension may
  reflect a shared low-density phenomenology, testable with current and
  forthcoming observations.
\end{abstract}
