\documentclass[universe,article,submit,pdftex,moreauthors]{Definitions/mdpi}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{physics}
\usepackage{tikz}
\usepackage{amsopn}
\usepackage{amstext}
\usepackage{microtype}
\usepackage{xparse}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
% Boxes
\usepackage[most]{tcolorbox}
% Plots (pgfplots provides axis/addplot)
\usepackage{pgfplots}
\usetikzlibrary{intersections}
\usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=1.18}
\NewDocumentCommand{\subsubsubsection}{s m}{%
  \IfBooleanTF{#1}{\paragraph*{#2}}{\paragraph{#2}}%
}
\usetikzlibrary
{arrows.meta, calc, fit, decorations.pathmorphing, decorations.pathreplacing, shapes.geometric, shadings, positioning,
  decorations.markings, patterns}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{definition}

\firstpage{1}
\setcounter{page}{\@firstpage}
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2026}
\copyrightyear{2026}
\datereceived{ }
\daterevised{ } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
\doinum{}

\Title{A Single Effective Framework Underlying Galaxy Rotation Curves and the Hubble Tension}

\newcommand{\orcidauthorA}{0009-0001-7697-7868}

\Author{Jerome Beau$^{1,*}$}

\AuthorNames{Jerome Beau}

\address{%
  \quad Independent Researcher, France; jerome.beau@cosmochrony.org}

\corres{Correspondence: jerome.beau@cosmochrony.org}

\abstract
{We investigate whether two long-standing observational anomalies---flat galaxy rotation curves and the Hubble constant
tension---may originate from a common effective mechanism operating in low-density environments. We introduce a minimal
phenomenological framework in which departures from Newtonian expectations arise from a bounded saturation of the
effective gravitational response at the level of kinematic inference,
  without invoking dark matter particles or modifying the background Friedmann expansion or early-universe physics. In
  the galactic regime,
  the MOND phenomenology emerges as the low-acceleration limit of this bounded-response framework rather than as a
  fundamental modification of gravity. The description reduces to standard Newtonian behavior in high-density
  environments and introduces a smooth, finite effective response in diffuse systems. Applied to disk galaxies,
  the framework reproduces the main features of observed rotation curves across a broad range of morphologies using
  baryonic matter alone,
  with stable parameter values and no halo-by-halo tuning. Explicit comparisons with representative galaxies from the
  SPARC sample show that the effective saturation scale correlates with observed surface-density-dependent trends. At
  cosmological scales,
  the same bounded-response structure predicts environment-dependent shifts in the locally inferred expansion rate. For
  typical void density contrasts,
  an explicit order-of-magnitude estimate demonstrates that the deep-saturation regime applies,
  yielding a fractional shift $\Delta H/H \simeq \delta_\rho f_{\max}$
  without introducing additional cosmological degrees of freedom or altering the global expansion history. Cosmic voids
  thus provide a natural setting for a systematic offset between local and global determinations of the Hubble constant.
  We examine robustness, degeneracies with standard astrophysical effects, and limitations of the effective description,
  and outline distinctive observational signatures,
  including environment-dependent redshift drift and subtle weak-lensing modifications in extreme low-acceleration
  regimes. These results suggest that galaxy rotation curves and the Hubble tension may reflect a shared low-density
  phenomenology testable with current and forthcoming observations.}

\keyword{cosmology; galaxy rotation curves; Hubble tension; modified gravity; voids}

\providecommand{\bibcommenthead}{}

\begin{document}

  \input{01-introduction/introduction}
  \input{02-effective_model/effective_model}
  \input{03-rotation_curves/rotation_curves}
  \input{04-voids_hubble/voids_hubble}
  \input{05-robustness/robustness}
  \input{06-predictions/predictions}
  \input{07-conclusion/conclusion}

  \appendix
  \input{A-cosmo_resolution_proof}

  \dataavailability{%
    The galaxy rotation-curve data used in this work are drawn from the publicly available
    SPARC database~\cite{Lelli2016SPARC}.
    The numerical code used to generate the galaxy rotation-curve figures and associated
    diagnostics is openly available and archived with a persistent identifier.
    The version used in this work is available via Zenodo at
    \url{https://doi.org/10.5281/zenodo.18462369}~\cite{CosmochronySimulation}.%
  }

  \acknowledgments
  {The author acknowledges the use of large language models as an editorial and analytical assistant during manuscript
  preparation, including help with phrasing alternatives, consistency checks, and structured rewriting.
  All scientific claims, modelling choices, numerical results, and interpretations are the sole responsibility of the
  author.}

  \conflictsofinterest{The author declares no conflicts of interest.}

  \reftitle{References}
  \bibliography{references}

\end{document}
