\subsection{Effective Expansion Rate in Voids}
  \label{subsec:void_effect}

  Cosmic voids represent the deepest and most extended low-density regions in the late Universe.
  Within voids, baryonic and total matter densities are strongly suppressed,
  and typical gravitational accelerations fall below the saturation scale $a_\star$
  over large volumes~\cite{Keenan2013KBC}.

  In this regime, the effective description operates close to its
  low-density saturation limit.
  The relation between local kinematics and the underlying background expansion
  is therefore modified at the level of inference rather than at the level of
  the dynamical Friedmann equations.
  To leading order, this effect can be captured by an environment-dependent
  renormalization of the inferred expansion rate,
  \begin{equation}
    H_{\mathrm{loc}}
    =
    H_{\mathrm{global}}
    \left[
      1
      +
      \delta_{\mathrm{eff}}
    \right],
    \quad \text{with} \quad
    \delta_{\mathrm{eff}}
    =
    \delta_\rho \,
    f\!\left(
         \frac{g_{\mathrm{void}}}{a_\star}
    \right),
    \label{eq:hloc}
  \end{equation}
  where
  $\delta_\rho = 1 - \rho_{\mathrm{void}}/\rho_{\mathrm{global}}$
  denotes the density contrast of the void,
  and $g_{\mathrm{void}}$ represents a characteristic gravitational
  acceleration scale associated with the void,
  estimated from its typical size and density contrast.

  The function $f(x)$ is a dimensionless saturation function
  determined by the same bounded-response structure that governs
  galactic dynamics.
  In the deep saturated regime ($x \ll 1$),
  $f(x)$ approaches a finite asymptotic value $f_{\max}$,
  set by the maximal projection deficit allowed by the effective description.
  Independent constraints derived in Section~4.3 of~\cite{Beau2026c}
  imply $f_{\max} \lesssim 0.1$.

  The saturation scale $a_\star$ entering Eq.~\eqref{eq:hloc}
  is fixed independently from galactic rotation curves
  (Section~\ref{subsec:universality}),
  and no cosmological parameter enters its determination.

  For void density contrasts in the observationally inferred range
  $\delta_\rho \sim 0.2$--$0.4$~\cite{Keenan2013KBC},
  and for $g_{\mathrm{void}} \ll a_\star$,
  Eq.~\eqref{eq:hloc} yields a characteristic fractional shift
  $\Delta H/H$ at the few-percent level.
  This magnitude is comparable to the observed local offset,
  without requiring any tuning of cosmological parameters.

  The magnitude of the effect therefore follows directly from combining
  (i) the empirically determined scale $a_\star$,
  and
  (ii) independently measured void density contrasts.

  The correction $\delta_{\mathrm{eff}}$ depends on the depth and spatial extent
  of the void, as well as on the empirically determined saturation scale
  $a_\star$.
  Denser or more weakly underdense regions remain close to the unsaturated regime
  and yield $\delta_{\mathrm{eff}} \approx 0$.

  Importantly, this mechanism does not introduce any additional energy component
  and does not modify the background Friedmann equations.
  The global expansion history remains unchanged.
  Only the local kinematic inference of the expansion rate is affected.

  Within this interpretation, the enhanced expansion inferred in cosmic voids
  reflects a reduced effective geometric constraint rather than an additional
  energy density.
  In regions where the density of contributing relational structure is low,
  the kinematic inference based on sparse mass distributions operates closer
  to its saturation limit, inducing a systematic environment-dependent bias
  in locally inferred kinematic relations.
