\subsection{Connection to the Hubble Tension}
  \label{subsec:hubble_tension}

  The Hubble tension refers to the discrepancy between early-time
  determinations of the Hubble constant, primarily inferred from
  cosmic microwave background measurements,
  and late-time local measurements based on the distance ladder~\cite{Riess2022SH0ES,Planck2020}.
  Despite extensive investigation, no single explanation has yet
  achieved broad consensus~\cite{Verde2019Review}.

  In the present framework, this discrepancy is interpreted
  as a manifestation of environment-dependent kinematic inference.
  Early-time measurements are anchored in the nearly homogeneous,
  high-density Universe, where the effective description operates
  in the unsaturated regime and recovers $H_{\mathrm{global}}$.
  By contrast, late-time distance-ladder measurements are performed
  in a highly structured Universe and may preferentially probe
  void-dominated environments.

  If observers and calibrator systems reside in regions operating
  near the saturated regime, Eq.~\eqref{eq:hloc} predicts a systematic
  upward shift in the locally inferred expansion rate,
  \begin{equation}
    H_{\mathrm{loc}} > H_{\mathrm{global}}.
  \end{equation}
  The magnitude of the shift depends on the depth and spatial
  extent of the surrounding underdensity.

  Importantly, this mechanism does not alter the background Friedmann evolution.
  It does not introduce new relativistic degrees of freedom,
  nor does it modify early-time cosmology.
  The effect arises solely from the bounded-response structure
  governing late-time low-density environments.

  The extent to which local underdensities alone can account
  for the full observed tension remains debated~\cite{Shanks2019Void,Kennworthy2019VoidCritique}.
  The present framework does not posit a single finely tuned large-scale void.
  Instead, it provides a distinct mechanism in which the same
  empirically determined saturation scale $a_\star$ that governs
  galactic dynamics induces a systematic environment-dependent
  bias in late-time kinematic inference.

  The proposal is therefore not a local-void cosmology,
  but a unified effective description in which both galaxy
  rotation curves and the Hubble offset arise from a common bounded-response principle.

  Because the dynamics derives from an effective potential $\Phi_{\mathrm{eff}}$,
  the framework admits a conserved effective energy and a
  corresponding modified virial relation for stationary systems.
  This ensures that equilibrium disk galaxies remain dynamically
  stable in the transition regime and that the void mechanism
  does not introduce inconsistencies at galactic scales.
