\subsection{Degeneracies with Standard Astrophysical Effects}
  \label{subsec:degeneracies}

  Any proposed alternative mechanism must be carefully examined against
  established astrophysical and cosmological effects within the
  $\Lambda$CDM framework.
  We therefore distinguish between degeneracies at galactic and cosmological scales.

  \paragraph{Galactic scales.}

    Within dark matter halo models, baryonic feedback processes and
    baryon--halo coupling have been shown to modify inner halo profiles
    and reproduce certain features of flattened rotation curves,
    particularly in low-mass systems~\cite{DiCintio2014Feedback}.
    Such models typically introduce galaxy-dependent halo parameters
    and rely on a dynamical coupling between baryonic surface density
    and halo response.

    By contrast, the effective saturation framework introduces no halo
    degrees of freedom and involves a single universal acceleration scale
    $a_\star$.
    All galaxy-to-galaxy diversity arises solely from the observed baryonic
    mass distribution and from the fitted stellar mass-to-light ratio.
    The resulting scaling relations follow directly from the structure of
    Eq.~\eqref{eq:modified_poisson}, without requiring halo-by-halo tuning
    or environment-dependent structural parameters.

    A degeneracy may therefore exist at the level of rotation-curve
    fitting accuracy.
    However, the underlying parameter structure differs qualitatively:
    halo models accommodate diversity through adjustable halo profiles,
    whereas the present framework attributes diversity to variations
    in baryonic structure under a fixed saturation scale.
    This difference implies distinct residual patterns and scaling
    correlations, providing a concrete avenue for empirical discrimination
    in detailed rotation-curve analyses.

  \paragraph{Cosmological scales.}

    At low redshift, local inhomogeneities, peculiar velocities,
    and sample variance are known to affect distance-ladder determinations
    of the Hubble constant~\cite{Freedman2021TRGB}.
    Large-scale underdensities have also been explored as partial
    explanations of the Hubble tension~\cite{Shanks2019Void,Kennworthy2019VoidCritique}.
    In most such analyses, these effects are treated as perturbations
    of the background expansion within standard gravity.

    The present framework differs in a crucial respect.
    The background Friedmann expansion remains unmodified.
    The predicted offset arises not from a change in global dynamics,
    nor from a radially tuned LTB-like void profile,
    but from an environment-dependent modification of effective
    kinematic inference once the local acceleration falls below
    the saturation scale $a_\star$.

    In particular, the mechanism does not rely on a large violation
    of homogeneity or on fine-tuned radial density profiles.
    Instead, it predicts a systematic scaling of the inferred expansion rate
    with the local density contrast, as encoded in Eq.~\eqref{eq:hloc}.
    This scaling is tied to the same universal parameter $a_\star$
    determined independently from galaxy rotation curves.

  \paragraph{Conceptual distinction.}

    A central ingredient of the framework is the existence of a universal
    bound on the effective response of the geometric description.
    This bound is not environment dependent.
    What varies across environments is the density of relations contributing
    to the effective projection.

    Dense systems concentrate many effective relations within limited volumes,
    keeping the dynamics close to the unsaturated regime.
    Extended low-density regions distribute the same fixed descriptive
    resolution over much larger volumes, pushing the effective description
    closer to saturation.
    The resulting enhancement of the locally inferred expansion rate
    therefore reflects how different environments sample a fixed
    bounded-response structure, rather than a modification of gravity,
    a change in the Friedmann background, or a reparameterization
    of dark-sector physics.

    This structural distinction cannot be absorbed into standard
    astrophysical degeneracies.
    It predicts correlated galactic and cosmological signatures tied
    to a single universal scale, providing a clear and testable
    avenue for empirical discrimination.
