\subsection{Little Red Dots as Early Saturation Signatures}
  \label{subsec:lrd_saturation}

  The bounded-response structure discussed throughout this work is not
  restricted to late-time, low-density environments.
  It also imposes consistency constraints in high-curvature primordial regimes.

  Recent JWST observations have revealed a population of compact,
  luminous, high-redshift sources often referred to as ``Little Red Dots''
  (LRDs)~\cite{Pacucci2025,Rusakov2026}.
  These objects have been interpreted as rapidly growing massive black holes,
  possibly embedded in dense ionized cocoons, implying efficient early
  mass assembly at redshifts $z \gtrsim 6$.

  Within the present framework, such observations do not require
  modifications of early-universe physics.
  They are structurally compatible with the same bounded-response principle
  if primordial systems approach the maximal admissible projective capacity
  under extreme compression.

  Two non-exclusive realizations are possible.
  First, halos exceeding a critical mass $M_{\mathrm crit}(z)$ may undergo
  genuine direct collapse, producing massive compact seeds at early epochs.
  Second, a strongly constrained relational configuration may enhance the
  effective convergence through the emergent contribution
  $\kappa_{\chi}$ introduced in Section~11.10 of the White Paper,
  without requiring proportionally large baryonic mass.

  In both cases, the mechanism is identical at the effective level:
  a resolution gap between the demanded fragmentation scale and the
  minimal projectable scale~$\ell(z)$.
  The observational signature is therefore not necessarily the presence
  of an exotic component, but the structural response of the inference
  metric near its saturation threshold.

  This interpretation is introduced as a consistency check rather than
  a standalone prediction.
  The framework does not claim that LRDs uniquely validate the mechanism,
  but that their inferred mass discrepancy does not contradict the same
  bounded-response structure that governs the low-density Hubble
  inference discussed above.

  Discriminating between a physical direct-collapse black hole and a
  predominantly projective overmass configuration requires combined
  constraints from variability, spectral diagnostics, and dynamical
  mass estimates.
  The framework predicts that the abundance of such compact early
  objects should track the redshift evolution of the effective
  resolution scale rather than solely the mass accretion history of halos.
